---
title: "Automated Machine Learning with H2O I"
author: "Mohamed Eltabei"
date: "2023-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Preperation

```{r}
# Libraries 
library(tidyverse)
library(readxl)
library(skimr)
library(GGally)

# Load Data data definitions
employee_attrition_tbl <- read_excel("./../../assets/DataSets/datasets-1067-1925-WA_Fn-UseC_-HR-Employee-Attrition.xlsx")
path_data_definitions <- "./../../assets/DataSets/data_definitions.xlsx"
definitions_raw_tbl   <- read_excel(path_data_definitions, sheet = 1, col_names = FALSE)



# Descriptive Features
employee_attrition_tbl %>% select(Age, DistanceFromHome, Gender, MaritalStatus, NumCompaniesWorked, Over18)

# Employment Features
employee_attrition_tbl %>% select(Department, EmployeeCount, EmployeeNumber, JobInvolvement, JobLevel, JobRole, JobSatisfaction)

# Compensation Features
employee_attrition_tbl %>% select(DailyRate, HourlyRate, MonthlyIncome, MonthlyRate, PercentSalaryHike, StockOptionLevel)

# Survery Results
employee_attrition_tbl %>% select(EnvironmentSatisfaction, JobSatisfaction, RelationshipSatisfaction, WorkLifeBalance)

# Performance Data
employee_attrition_tbl %>% select(JobInvolvement, PerformanceRating)

# Work-Life Features
employee_attrition_tbl %>% select(BusinessTravel, OverTime)

# Training & Education
employee_attrition_tbl %>% select(Education, EducationField, TrainingTimesLastYear)

# Time-Based Features
employee_attrition_tbl %>% select(TotalWorkingYears, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager)




# Step 1: Data Summarization -----

skim(employee_attrition_tbl)

# Character Data Type
employee_attrition_tbl %>%
  select_if(is.character) %>%
  glimpse()

# Get "levels"
employee_attrition_tbl %>%
  select_if(is.character) %>%
  map(unique)

# Proportions    
employee_attrition_tbl %>%
  select_if(is.character) %>%
  map(~ table(.) %>% prop.table())

# Numeric Data
employee_attrition_tbl %>%
  select_if(is.numeric) %>%
  map(~ unique(.) %>% length())

employee_attrition_tbl %>%
  select_if(is.numeric) %>%
  map_df(~ unique(.) %>% length()) %>%
  # Select all columns
  pivot_longer(everything()) %>%
  arrange(value) %>%
  filter(value <= 10)


employee_attrition_tbl %>%
  select(Attrition, Age, Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome) %>%
  ggpairs() 


employee_attrition_tbl %>%
  select(Attrition, Age, Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome) %>%
  ggpairs(aes(color = Attrition), lower = "blank", legend = 1,
          diag  = list(continuous = wrap("densityDiag", alpha = 0.5))) +
  theme(legend.position = "bottom")



# Create data tibble, to potentially debug the plot_ggpairs function (because it has a data argument)
data <- employee_attrition_tbl %>%
  select(Attrition, Age, Gender, MaritalStatus, NumCompaniesWorked, Over18, DistanceFromHome)

plot_ggpairs <- function(data, color = NULL, density_alpha = 0.5) {
  
  color_expr <- enquo(color)
  
  if (rlang::quo_is_null(color_expr)) {
    
    g <- data %>%
      ggpairs(lower = "blank") 
    
  } else {
    
    color_name <- quo_name(color_expr)
    
    g <- data %>%
      ggpairs(mapping = aes_string(color = color_name), 
              lower = "blank", legend = 1,
              diag = list(continuous = wrap("densityDiag", 
                                            alpha = density_alpha))) +
      theme(legend.position = "bottom")
  }
  
  return(g)
  
}
```
# Analysis

## Compensation Features

```{r}
employee_attrition_tbl %>%
  select(Attrition, contains("income"), contains("rate"), contains("salary"), contains("stock")) %>%
  plot_ggpairs(Attrition)

```
1. What can you deduce about the interaction between Monthly Income and Attrition?

Answer: That those are staying have a lower Monthly Income.

2. What can you deduce about the interaction between Percent Salary Hike and Attrition?

Answer: It's difficult to deduce anything based on the visualization

3. What can you deduce about the interaction between Stock Option Level and Attrition?

Answer: Those that are leaving the company have a higher stock option level

## Survey Results: Satisfaction level, WorkLifeBalance 


```{r}

employee_attrition_tbl %>%
    select(Attrition, contains("satisfaction"), contains("life")) %>%
    plot_ggpairs(Attrition)

```

4. What can you deduce about the interaction between Environment Satisfaction and Attrition?

Answer: A higher proportion of those leaving have a high environment satisfaction level

5. What can you deduce about the interaction between Work Life Balance and Attrition?

Answer: Those that are staying have a lower density of 2's and 3's

## Performance Data: Job Involvment, Performance Rating

```{r}
employee_attrition_tbl %>%
    select(Attrition, contains("performance"), contains("involvement")) %>%
    plot_ggpairs(Attrition)
```

6. What Can you deduce about the interaction between Job Involvement and Attrition?

Answer: Those that are staying have a lower density of 2's and 3's

## Work-Life Features

```{r}
employee_attrition_tbl %>%
  select(Attrition, contains("overtime"), contains("travel")) %>%
  plot_ggpairs(Attrition)

```

7. What can you deduce about the interaction between Over Time and Attrition?

Answer: The proportion of those leaving that are working Over Time are high compared to those that are not leaving

## Training and Education

```{r}

employee_attrition_tbl %>%
  select(Attrition, contains("training"), contains("education")) %>%
  plot_ggpairs(Attrition)

```
8. What can you deduce about the interaction between Training Times Last Year and Attrition?

Answer: People that leave tend to have more annual training

## Time-Based Features: Years at company, years in current role

```{r}

employee_attrition_tbl %>%
  select(Attrition, contains("years")) %>%
  plot_ggpairs(Attrition)

```

9. What can you deduce about the interaction between Years At Company and Attrition?

Answer: People that leave tend to have more working years at the company

10. What can you deduce about the interaction between Years Since Last Promotion and Attrition?

Answer: It's difficult to deduce anything based on the visualization
